Physics-informed Implicit Neural Representation for Potential Field Geophysics

### 2023-08-19 paper3.tex
Add results text

\section{Results}
\subsection{Implicit gridding}
\begin{figure}
    \label{fig:grid}
    \caption[]{Constant level grid predicted at the survey nominal terrain clearance \SI{40}{m}}
\end{figure}

\begin{enumerate}
    \item{} Best attempt at recreating grid as gridded by GA\@.
    \item{} Demonstration of the range of altitude slices
\end{enumerate}

\subsection{Spatial gradients and filters}
\begin{enumerate}
    \item{} Horizontal Gradients
    \item{} Vertical Gradients
    \item{} Combinations and filters
\end{enumerate}

\section{Discussion}
\begin{enumerate}
    \item{} Limitations: Capacity of Network
    \item{} RAM usage (if I can't do batched C-MLP training)
    \item{} Accuracy of learnt x,y and z gradients
    \item{} Forward looking at the field of INR
\end{enumerate}

% With this representation the extraction of regularised two dimensional slices, namely grids, could be performed \emph{a posteri} of gridding.
% Regularising these samples to a two dimensional grid for interpretation, storage, and further machine processing is a well studied process with many methods available.
% After regularisation, the grid arrays can be subject to a large number of widely used filters, such as horizontal and vertical gradient calculation and their derivatives.
% These filters operate on the grid product, a representation of the potential field quantised by the cell spacing defined during the regularisation process.

Implicit neural representation is a rapidly evolving field of research with many advances occuring within each year, borrowing contemporary deep learning improvements from related fields.
In addition to constant work on improving the high-frequency learning capacity, the inclusion of additional dimensions or

\section{Conclusions}

>>>>

\section{Results}
\subsection{Implicit gridding}
Querying a learnt implicit neural represenation (INR) using a regularly spaced \(x,y\) grid at a specific \(z\) height is a novel method for gridding geophysical potential field surveys.
\Cref{fig:grid} shows the INR queried at \(z=\SI{40}{\meter}\), which is the nominal terrain clearance of the survey.
Notably, the cell size can be selected by fining or coarsening the \(x,y\) spacing, including variable or anisotropic cell sizes.
For natural images, using finer cell sizes accomplishes the computer vision task of super-resolution.
However, for the potential field representation task, no additional high-frequency information is predicted by the MLP network and resolution remains limited by the survey sample spacing and neural network capacity.

\begin{figure}[hbt]
    \includegraphics[width=\textwidth]{fig/p3/grid_comparison.pdf}
    \caption[Grid prediction]{Comparison between the ground-truth grid processed by Geoscience Australia (left) and the constant level grid queried from the INR at \SI{40}{\m} (right).}
    \label{fig:grid}
\end{figure}

The INR can be queried at any continuous coordinate, including out-of-range values beyond the training domain.
The predicted field becomes unstable beyond this boundary, which can be seen in the periphery of \cref{fig:grid}.

Constant level grids can be queried from the INR at any height.
\Cref{fig:grid35} shows the case study extent queried at \SI{35}{m}.
While comparable features can still be interpreted within this level grid, the represenation learnt here fails to represent physically realistic continuation of the field.

\begin{figure}[hbt]
    %TODO query at AWAGS height and compare GA AWAGS grid.
    % TODO check if provided grid is AWAGS or MICRO levelled - maybe grid data myself.
    \includegraphics[width=\textwidth]{fig/p3/grid_comparison_35m.pdf}
    \caption[Grid prediction]{A constant level grid queried from the INR at the AWAGS datum height of \SI{35}{m} (right).}
    \label{fig:grid35}
\end{figure}

% The strong performance of the model is retained even when using a subsampled training dataset, here shown at \qty{30}{\percent} and \qty{10}{\percent} (\cref{fig:gridpct}) of the total available data.
% % In both cases, the validaiton and test sets remain identical to the initial training split.

% \begin{figure}[hbt]
%     % \includegraphics[width=\textwidth]{fig/p3/grid_subpct.pdf}
%     \caption[Grid subsampled]{Performance of level grid prediction shown at \qty{30}{\percent} and \qty{10}{\percent} random sampling of training data.}
%     \label{fig:gridpct}
% \end{figure}

\subsection{Spatial derivatives and filters}
Using AD, derivatives can be calculated between the input spatial coordinates and the output potential field represenation.
While the horizontal derivatives correspond well to numerical derivatives calculated on the Wolfe Creek ground truth grid (\cref{fig:hori_grad}), the vertical derivative suffers poor performance (\cref{fig:vert_grad}).
This correlates with the poor continuation performance of model when predicting level grids several meters or more beyond the nominal sample height.
This is inrepreted as the model learning a poor represenation of the vertical component of the potential field.

\begin{figure}[hbt]
    % \includegraphics[width=\textwidth]{fig/p3/hori_grad.pdf}
    \caption[Horizontal derivatives]{AD calculated horizontal derivatives of the INR\@. Left: Northing derivative, right: easting derivative}
    \label{fig:hori_grad}
\end{figure}

\begin{figure}[hbt]
    % \includegraphics[width=\textwidth]{fig/p3/vert_grad.pdf}
    \caption[Vertical derivative]{AD calculated vertical derivative of the INR.}
    \label{fig:vert_grad}
\end{figure}

Using the implicit representation, a selection of common geophysical filters are calculated prior to quantisation.

\begin{enumerate}
    \item ASIG
    \item tilt?
    \item idk?
\end{enumerate}

\section{Discussion}
The implicit represenation has strong representive capacity for the horizontal components of the potential field, generating level grids which are accurate against the ground truth grid.
Level grids can be generated at any height, with best performance achieved near the realised nominal survey altitude.
Performance worsens within approximately one standard deviation from that height (\SI{35} to \SI{45}{\m}), with gradual failure of the model outisde of that height range.
% Even when using a small subset of the dataset, this performance is retained.
It is interpreted from the poor quality of the vertical gradients and out-of-range predictions that the model is relying on learning a function between horizontally coplanar points, as opposed to a realistic vertical continuation function.
While quantised grids can be queried at any cell size, short wavelength information beyond the Nyquist limit of the survey line sampling is not predicted.
% TODO The addition of a PINN term for objective function ...

The case study survey presented is very small compared to typical regional aeromagnetic surveys.
A demonstration of the method on a large aeromagnetic survey, requiring minibatch training, is included in \cref{app:menzies}.
Training of this representation follows the parameters outlined in \cref{sec:training}, which were optimised for the Wolfe Creek case study, with the addition of a shuffled minibatch size of \SI{1024000}.
Attempts to train the CMLP with small batch sizes did not converge.%, reflecting the non-convergance result found for training on very large subsampling factors of the line dataset.

Implicit neural representation is a rapidly evolving field of research, and draws on contemporary deep learning improvements within related fields.
In addition to improving the high-frequency learning capacity, physics informed loss functions promise higher quality vertical continuation performance. 
The current method could trivially include additional target dimensions, which would allow representation learning of vector fields \(\mathbb{R}^{3}\) in three dimensions \(\mathbb{R}^{3}\).
This has potential application in tensor fields and associated predictive tasks.
These remain for future work.

\section{Conclusions}
High-resolution implicit neural represenation of potential field geophysics surveys is achieved with the use of coordinate MLP neural networks.
Resolution in the represenation is limited by network capacity, rather than initial grid cell size selection during regularisation.
Gradients and spatial filters can be caluclated using the full resolution of the learnt represenation, and gridded \emph{a posteri} of gradient or filter calculation.
